### 2008-09-05 trunk/Papers/rmrs-08/abstract.tex
language more concise and partial conclusion added

\begin{abstract}
  la la la
\end{abstract}

>>>>

\begin{abstract}
  One way to construct semantic
  representations in a robust manner is to enhance shallow language
  processors with semantic components.  Here, we provide a model
  theory for a 
  semantic formalism 
  that is designed for this,
  namely Robust Minimal Recursion Semantics ({\sc rmrs},
  \cite{copestake:2003}).  We show that {\sc rmrs} supports a notion
  of entailment 
  that allows it
  to form the basis for comparing the semantic output of
  different parses of varying depth.
\end{abstract}

### 2008-09-05 trunk/Papers/rmrs-08/conclusion.tex
first dummy version
language more concise and partial conclusion added

\begin{abstract}
  la la la
\end{abstract}

>>>>

\section{Conclusion}
\label{sec:conclusion}

In this paper we have motivated and described \rmrs---a semantic
framework that can be used to specify
semantic components to robust shallow parsers.  Accordingly, \rmrs\
is designed to be
maximally flexible on the type of semantic information
that can be left underspecified, so that the semantic output of a
shallow parser needn't over-determine or under-determine the semantic
information that can be extracted from the shallow syntactic analysis.

We defined its syntax and model-theoretic semantics, and following the
tradition from dominance constraints, we defined a syntactic criterion
for characterising entailment.  The syntactic notion of {\em solved
  forms} is useful because it means that even when an underspecified
\rmrs\ corresonds to an infinite number of possible fully-specific
interpretations in context, there is a finite way of identifying all
the `relevant' differences among those fully-specific interpretations.
We showed that the model theory supports theorems that allow one to
use \rmrs\ representations as a way of comparing and even integrating
the semantic output of different parses of the same natural language
string, even when they vary in depth.

### 2008-09-05 trunk/Papers/rmrs-08/introduction.tex
language more concise and partial conclusion added

Representing semantics as a logical form that supports automated
inference and model construction is vital for deeper language
engineering tasks, such as dialogue systems.  Such detailed semantic
representations can be obtained from hand-crafted deep grammars (e.g.,
\cite{butt:etal:1999,copestake:flickinger:2000}), but this
tends to lack robustness: hand-crafted grammars fail to cover all
words and linguistic constructions; and unedited text and speech
contains ill-formed phrases, which by design deep grammars do not
handle.

On the other hand, there has been a recent trend towards systems for
robust wide-coverage semantic construction.  For instance,
\cite{bos:etal:2004} exploit
{\sc ccg}'s elegant syntax-semantics interface to produce a unique
logical form for each derivation of
a statistical {\sc ccg} parser;
\cite{zettlemoyer:collins:2007} and others learn
string-to-logical-form mappings, exploting informative features from
the output of
shallow language processors; and \cite{copestake:2003} and
\cite{frank:2004}, among others,
endow robust language processors with semantic components.

However, there are certain semantic
phenomena that these robust
approaches don't capture reliably, including quantifier scope, optional
arguments,
% COMMENT: this issue about optional arguments is really too cryptic.
% Basically, there is no way for assessing by *any* learned language
% processor whether a single word form that appears in
% a number of different syntactic valencies is an example of lexical
% sense ambiguity, or they are all examples of the same sense being
% used in syntactically different ways (i.e., we have optional
% arguments).  Only hand-crafted grammars can tell you this.  So, for
% example, Bos et al (2004) would get bet(x,y) for "I bet 5 pounds",
% and bet(x,y,z) for "I bet you 5 pounds" and bet(x,p) for "I bet it's
% going to rain": I.e., the vocab of the semantic formalism is
% constructed automatically from NL word forms, and the predicate
% symbols in that automatically constructed vocab vary in (a) valency,
% and (b) the semantic dependencies represented by a given argument
% position (e.g., the second argument of "bet" in "I bet 5 pounds" is
% 5 pounds, and for "I bet it's raining" it's "it's raining".
% 
and long-distance dependencies (for instance, \cite{clark:etal:2004}
report that the parser used in \cite{bos:etal:2004} yields semantic
dependencies from the unbounded dependencies borne from object
extraction---e.g., {\em the man that I met\dots}---with 63\%
accuracy).  Forcing a shallow parser to make a decision about these
phenomena therefore can be a major source of errors.  Depending on the
application, it may be preferrable to give the parser the option to to
leave the decision open when it's not sufficiently informed---i.e., to
compute a partial semantic representation that neither over-determines
nor under-determines the semantic information revealed by the
(shallow) syntactic analysis, and to complete the semantic information
later, via inferences that exploit information extraneous to the
parser.

In this paper, we focus on one particular approach to representing
partial semantic information:
Robust Minimal Recursion Semantics ({\sc rmrs},
\cite{copestake:2003}).  {\sc rmrs} is a generalisation of {\sc mrs}
\cite{copestake:etal:2005} that is designed to support underspecification both
of scope relations and of the predicate-argument structure.
\cite{copestake:2003,frank:2004} use it to
support semantic constructions by shallow parsers
ranging in depth from {\sc pos} taggers to chunk parsers  
and intermediate parsers such as {\sc rasp}
\cite{briscoe:etal:2006}.  {\sc mrs} analyses
derived from deep grammars, such as the English Resource Grammar
\cite{copestake:flickinger:2000} can also be seen as special cases of
{\sc rmrs} representations.

The key contribution we make is to cast {\sc rmrs}, for the first
time, as a logic with a clearly defined model theory.  This has a
number of advantages.  First, it allows us to formally relate \rmrs\ to
other frameworks such as \mrs\ and predicate logic in order to
establish the compatibility of \rmrs\ with the approaches to robust
semantic analysis mentioned above.  Second, it provides us with a
notion of logical entailment, which we can use to verify the
compatibility of \rmrs\ representations computed from different
parsers, such as a deep and a shallow parser.  Finally, it is only
through rigid formalisation that we can give users of \rmrs\ access to
inference systems and to algorithms for efficiently computing the
possible fully specific semantic representations described by the \rmrs
output of a shallow parser.

We will first illustrate the differences and challenges with deep and
shallow semantic construction in an example in
Section~\ref{sec:motivation}.  We will then define the syntax and
model-theoretic semantics of \rmrs\ and introduce some important
formal concepts in Section~\ref{sec:rmrs}.  Finally, we demonstrate
the practical usefulness of our formalisation in
Section~\ref{sec:entailment} by modelling compatibility of
partial semantic representations as \rmrs\ entailment and providing a
syntactic criterion for characterising entailment.
Section~\ref{sec:conclusion} concludes and points to future work.

\hidden{

The language in which these partial representations are expressed

>>>>

Representing semantics as a logical form that supports automated
inference and model construction is vital for deeper language
engineering tasks, such as dialogue systems.  Logical forms
can be obtained from hand-crafted deep grammars
\cite{butt:etal:1999,copestake:flickinger:2000}, but this
tends to lack robustness: hand-crafted grammars fail to cover all
words and linguistic constructions and by design they do not handle
ill-formed phrases.

On the other hand, there has been a recent trend towards systems for
robust wide-coverage semantic construction.  For instance,
\cite{bos:etal:2004} exploit
{\sc ccg}'s elegant syntax-semantics interface to produce a unique
logical form for each derivation of
a statistical {\sc ccg} parser;
\cite{zettlemoyer:collins:2007} and others learn
string-to-logical-form mappings, exploting informative features from
shallow language processors; and \cite{copestake:2003} and
\cite{frank:2004}
endow robust language processors with semantic components.

However, there are certain semantic
phenomena that these robust
approaches don't capture reliably, including quantifier scope, optional
arguments,
% COMMENT: this issue about optional arguments is really too cryptic.
% Basically, there is no way for assessing by *any* learned language
% processor whether a single word form that appears in
% a number of different syntactic valencies is an example of lexical
% sense ambiguity, or they are all examples of the same sense being
% used in syntactically different ways (i.e., we have optional
% arguments).  Only hand-crafted grammars can tell you this.  So, for
% example, Bos et al (2004) would get bet(x,y) for "I bet 5 pounds",
% and bet(x,y,z) for "I bet you 5 pounds" and bet(x,p) for "I bet it's
% going to rain": I.e., the vocab of the semantic formalism is
% constructed automatically from NL word forms, and the predicate
% symbols in that automatically constructed vocab vary in (a) valency,
% and (b) the semantic dependencies represented by a given argument
% position (e.g., the second argument of "bet" in "I bet 5 pounds" is
% 5 pounds, and for "I bet it's raining" it's "it's raining".
% 
and long-distance dependencies (for instance, \cite{clark:etal:2004}
report that the parser used in \cite{bos:etal:2004} yields 63\%
accuracy on object
extraction; e.g., {\em the man that I met\dots}).
Forcing a shallow parser to make a decision about these
phenomena can therefore be error-prone.  Depending on the
application, it may be preferrable to give the parser the option to to
leave the decision open when it's not sufficiently informed---i.e., to
compute a partial semantic representation that neither over-determines
nor under-determines the semantic information revealed by the
(shallow) syntactic analysis, and to complete the semantic information
later, using information extraneous to the
parser.

In this paper, we focus on one particular approach to representing
partial semantic information:
Robust Minimal Recursion Semantics ({\sc rmrs},
\cite{copestake:2003}).  {\sc rmrs} is a generalisation of {\sc mrs}
\cite{copestake:etal:2005} that is designed to support underspecification
of scope and predicate-argument structure.
\cite{copestake:2003,frank:2004} use it to
specify semantic components to shallow parsers
ranging in depth from {\sc pos} taggers to chunk parsers  
and intermediate parsers such as {\sc rasp}
\cite{briscoe:etal:2006}.  {\sc mrs} analyses
derived from deep grammars, such as the English Resource Grammar
({\sc erg}, \cite{copestake:flickinger:2000}) are
special cases of 
{\sc rmrs}.

The key contribution we make is to cast {\sc rmrs}, for the first
time, as a logic with a clearly defined model theory.  This has a
number of advantages.  First, it allows us to formally relate \rmrs\ to
other frameworks such as \mrs\ and predicate logic in order to
establish the compatibility of \rmrs\ with the approaches to robust
semantic analysis mentioned above.  Second, it provides us with a
notion of logical entailment, which we can use to verify the
compatibility of \rmrs\ representations computed from different
parsers, such as a deep and a shallow parser.  Finally, it is only
through rigid formalisation that we can give users of \rmrs\ access to
inference systems and to algorithms for efficiently computing the
possible fully specific semantic representations described by the \rmrs\
output of a shallow parser.

Section~\ref{sec:motivation} illustrates via example the differences
and challenges with deep and 
shallow semantic construction.  
Section~\ref{sec:rmrs} defines the syntax and
model-theory of \rmrs\.  In Section~\ref{sec:entailment} we
demonstrate 
the practical usefulness of our formalisation 
by modelling compatibility of
partial semantic representations as \rmrs\ entailment and by providing a
syntactic criterion for characterising entailment.
Section~\ref{sec:conclusion} concludes and points to future work.

\hidden{
The language in which these partial representations are expressed

### 2008-09-05 trunk/Papers/rmrs-08/motivation.tex
language more concise and partial conclusion added

We motivate and describe the main features of {\sc rmrs} with a toy
example.  Sentence 
(\ref{eg:cat}a) features a semantic scope ambiguity that, according to
most deep grammars, syntax
doesn't resolve.    The {\sc mrs} (\ref{eg:cat}b) derived by the
English Resource Grammar ({\sc erg}, \cite{copestake:flickinger:2000})
reflects this: (i) each predication is
labelled ($l_1$, $l_4$, etc); (ii) scopal arguments are indicated by
holes (e.g., $h_2$ and $h_3$ in the quantifier {\em every}); (iii)
constraints on scope are expressed by the qeq conditions (roughly, $h
=_q l$ means that only zero or more quantifiers lie between the scopal
argument $h$ and the predication labelled by $l$ in any fully-resolved
logical form); and (iv) these scope
constraints allow for more than one way of `plugging' the holes with
labels.\footnote{The event variable in the predicate symbol
  corresponding to the adjective is motivated by copulas---{\em The
    cat is fat} has logical form 
  $\mbox{\_the\_q\_1}(x,\mbox{\_cat\_n\_1}(x),\mbox{\_fat\_j\_1}(e,x))$.}
\begin{examples}
\item   \label{eg:cat}
\begin{subexamples}
\item
Every fat cat chased some dog.
% \item
% $\mbox{\_every\_q\_1}(x, \mbox{\_fat\_j\_1}(e',x) \wedge
% \mbox{\_cat\_n\_1}(x),$\\
% \hspace*{0.5in}$\mbox{\_some\_q\_1}(y, \mbox{\_dog\_n\_1}(y),
% \mbox{\_chase\_v\_1}(e_{spast},x) )$\\
% $\mbox{\_some\_q\_1}(y, \mbox{\_dog\_n\_1}(y),$\\
% \hspace*{0.5in}$\mbox{\_every\_q\_1}(x, \mbox{\_fat\_j\_1}(e',x)
% \wedge 
% \mbox{\_cat\_n\_1}(x), 
% \mbox{\_chase\_v\_1}(e_{spast},x)))$
\item   $l_1\handel\mbox{\_every\_q\_1}(x,h_2,h_3)$\\
$l_4\handel\mbox{\_fat\_j\_1}(e',x)$\\
$l_4\handel\mbox{\_cat\_n\_1}(x)$\\
$l_5\handel\mbox{\_chased\_n\_1}(e,x,y)$\\
$l_6\handel\mbox{\_some\_n\_1}(y,h_7,h_8)$\\
$l_9\handel\mbox{\_dog\_n\_1}(y)$\\
$h_2 =_q l_4, h_7 =_q l_9$
\end{subexamples}
\end{examples}
(\ref{eg:cat}b) also exhibits {\sc mrs}' compact description of
conjunction through label sharing (see $l_4$ above), and the {\sc
  erg}'s naming convention for predicate symbols: each predicate is
constructed from the word lemma, its syntactic category and a sense
number, and the leading underscore identifies lexical predicates.

@@

The {\sc mrs} (\ref{eg:cat}c) is a notational variant of the {\sc
  rmrs} (\ref{eg:cat}d). {\sc rmrs}s offer a more factorised
representation: its predicate symbols are all unary and the other
arguments to the (base-language) constructor that it corresponds to
are represented by separate binary relations on the unique {\em
  anchor} of the predicate symbol ($a_1, a_{41},\ldots$) together with
variable and label equalities (e.g., $x=x_1$, $l_{41}=l_{42}$).
\begin{examples}
\item   [\ref{eg:cat}]
\begin{subexamples}
\item   [c]
$l_1:a_1\handel\mbox{\_every\_q\_1}(x_1),$\\
\hspace*{0.1in}$\mbox{RSTR}(a_1,h_2),
\mbox{BODY}(a_1,h_3)$\\ 
$l_{41}:a_{41}\handel\mbox{\_fat\_j\_1}(e'), \mbox{ARG1}(a_{41},x_2)$\\
$l_{42}:a_{42}\handel\mbox{\_cat\_n\_1}(x_3)$\\
$l_5:a_5\handel\mbox{\_chase\_v\_1}(e_{spast})$,\\
\hspace*{0.1in}$\mbox{ARG1}(a_5,x_4),
\mbox{ARG2}(a_5,x_5)$\\ 
$l_6:a_6\handel\mbox{\_some\_q\_1}(x_6)$,\\
\hspace*{0.1in}$\mbox{RSTR}(a_6,h_7),
\mbox{BODY}(a_6,h_8)$\\ 
$l_9:a_9\handel\mbox{\_dog\_n\_1}(x_7)$\\
$h_2=_q l_{42}, l_{41}=l_{42}, h_7 =_q l_9$\\
$x_1=x_2, x_2=x_3, x_3=x_4,$\\
$x_5=x_6, x_5=x_7$
\end{subexamples}
\end{examples}

This factored representation allows one to build semantic components
to shallow parsers where lexical or syntactic information is absent.
An extreme example would be a {\sc 
  pos} tagger: one can build its semantic component simply by deriving
lexical predicate symbols from the word lemmas and their tags and ensuring
that the predication for each word receives a unique label, anchor and
argument, as given in (\ref{eg:cat-pos}):
\begin{examples}
\item   \label{eg:cat-pos}
\begin{subexamples}
\item   Every\_{\sc at1} fat\_{\sc jj} cat\_{\sc nn1} chased\_{\sc
    vvd} some\_{\sc dd} dog\_{\sc nn1}
\item   $l_1:a_1:\mbox{\_every\_q}(x)$, \\
$l_{41}:a_{41}:\mbox{\_fat\_a}(e')$,\\
$l_{42}:a_{42}:\mbox{\_cat\_n}(x_3)$\\
$l_5:a_5:\mbox{\_chased\_v}(e_{past})$, \\
$l_6:a_6:\mbox{\_some\_q}(x_4)$, \\
$l_9:a_9:\mbox{\_dog\_n}(x_5)$
\end{subexamples}
\end{examples}

\todo{The syntax used in this RMRS is incompatible with the later
  definition.}

>>>>

We motivate and describe {\sc rmrs} with a toy
example.  Sentence 
(\ref{eg:cat}a) features a semantic scope ambiguity that, according to
most deep grammars, syntax
doesn't resolve.    The {\sc mrs} (\ref{eg:cat}b) derived by the
{\sc erg} \cite{copestake:flickinger:2000}
reflects this: (i) each predication is
labelled ($l_1$, $l_4$, etc); (ii) scopal arguments are indicated by
holes (e.g., $h_2$, $h_3$); (iii)
constraints on scope are expressed by the qeq conditions (roughly, $h
=_q l$ means that only zero or more quantifiers lie between the scopal
argument $h$ and the predication labelled by $l$ in any fully-resolved
logical form); and (iv) these scope
constraints allow for more than one way of `plugging' the holes with
labels.\footnote{The event variable in $\_\mbox{{\em fat}}\_j\_1$ is 
  for copulas. {\em The
    cat is fat} has logical form 
  $\mbox{\_the\_q\_1}(x,\mbox{\_cat\_n\_1}(x),\mbox{\_fat\_j\_1}(e,x))$.}
\begin{examples}
\item   \label{eg:cat}
\begin{subexamples}
\item
Every fat cat chased some dog.
% \item
% $\mbox{\_every\_q\_1}(x, \mbox{\_fat\_j\_1}(e',x) \wedge
% \mbox{\_cat\_n\_1}(x),$\\
% \hspace*{0.5in}$\mbox{\_some\_q\_1}(y, \mbox{\_dog\_n\_1}(y),
% \mbox{\_chase\_v\_1}(e_{spast},x) )$\\
% $\mbox{\_some\_q\_1}(y, \mbox{\_dog\_n\_1}(y),$\\
% \hspace*{0.5in}$\mbox{\_every\_q\_1}(x, \mbox{\_fat\_j\_1}(e',x)
% \wedge 
% \mbox{\_cat\_n\_1}(x), 
% \mbox{\_chase\_v\_1}(e_{spast},x)))$
\item   $l_1\handel\mbox{\_every\_q\_1}(x,h_2,h_3)$\\
$l_4\handel\mbox{\_fat\_j\_1}(e',x)$\\
$l_4\handel\mbox{\_cat\_n\_1}(x)$\\
$l_5\handel\mbox{\_chased\_n\_1}(e,x,y)$\\
$l_6\handel\mbox{\_some\_n\_1}(y,h_7,h_8)$\\
$l_9\handel\mbox{\_dog\_n\_1}(y)$\\
$h_2 =_q l_4, h_7 =_q l_9$
\end{subexamples}
\end{examples}
(\ref{eg:cat}b) also exhibits {\sc mrs}' compact description of
conjunction through label sharing (see $l_4$), and the {\sc
  erg}'s naming convention for predicate symbols: each lexical predicate is
constructed from the word lemma, its syntactic category and a sense
number, and the leading underscore distinguishes lexical predicates
from non-lexical ones.

@@

The {\sc mrs} (\ref{eg:cat}b) is a notational variant of the {\sc
  rmrs} (\ref{eg:cat}c). {\sc rmrs}s offer a more factorised
representation: each predicate symbol is unary and the other
arguments to its corresonding base-language constructor
are represented by separate binary relations on the unique {\em
  anchor} of the predicate symbol ($a_1, a_{41},\ldots$) together with
variable equalities (e.g., $x=x_1$).
\begin{examples}
\item   [\ref{eg:cat}]
\begin{subexamples}
\item   [c]
$l_1:a_1\handel\mbox{\_every\_q\_1}(x_1),$\\
\hspace*{0.1in}$\mbox{RSTR}(a_1,h_2),
\mbox{BODY}(a_1,h_3)$\\ 
$l_{41}:a_{41}\handel\mbox{\_fat\_j\_1}(e'), \mbox{ARG1}(a_{41},x_2)$\\
$l_{42}:a_{42}\handel\mbox{\_cat\_n\_1}(x_3)$\\
$l_5:a_5\handel\mbox{\_chase\_v\_1}(e_{spast})$,\\
\hspace*{0.1in}$\mbox{ARG1}(a_5,x_4),
\mbox{ARG2}(a_5,x_5)$\\ 
$l_6:a_6\handel\mbox{\_some\_q\_1}(x_6)$,\\
\hspace*{0.1in}$\mbox{RSTR}(a_6,h_7),
\mbox{BODY}(a_6,h_8)$\\ 
$l_9:a_9\handel\mbox{\_dog\_n\_1}(x_7)$\\
$h_2=_q l_{42}, l_{41}=l_{42}, h_7 =_q l_9$\\
$x_1=x_2, x_2=x_3, x_3=x_4,$\\
$x_5=x_6, x_5=x_7$
\end{subexamples}
\end{examples}

This factored representation allows one to build semantic components
to shallow parsers where lexical or syntactic information is absent.
An extreme example would be a {\sc 
  pos} tagger: one can build its semantic component simply by deriving
lexical predicate symbols from the word lemmas and tags and assigning each
predication a unique label, anchor and
argument, as in (\ref{eg:cat-pos}):
\begin{examples}
\item   \label{eg:cat-pos}
\begin{subexamples}
\item   Every\_{\sc at1} fat\_{\sc jj} cat\_{\sc nn1} chased\_{\sc
    vvd} some\_{\sc dd} dog\_{\sc nn1}
\item   $l_1:a_1:\mbox{\_every\_q}(x)$, \\
$l_{41}:a_{41}:\mbox{\_fat\_a}(e')$,\\
$l_{42}:a_{42}:\mbox{\_cat\_n}(x_3)$\\
$l_5:a_5:\mbox{\_chased\_v}(e_{past})$, \\
$l_6:a_6:\mbox{\_some\_q}(x_4)$, \\
$l_9:a_9:\mbox{\_dog\_n}(x_5)$
\end{subexamples}
\end{examples}
